---
title: "Dosresp"
author: "Reynaldi Ikhsan Kosasih"
date: "2/2/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summarized data about the relation between red meat consumption
## and diabetes risk in 12 prospective cohort

```{r results='hide'}
library(dosresmeta)
library(readxl)
redmeat <- read_excel("repository data.xlsx", sheet = "Clean")
View(redmeat)
```

## Fixed-effect dose-response model assuming linearity
```{r results='hold'}
lin.fixed <- dosresmeta(formula = logrr ~ dose, type = type, id = id,
	se = se, cases = cases, n = peryears,  data = redmeat, method = "fixed")
summary(lin.fixed)
predict(lin.fixed, delta = 50, exp = TRUE)
```
## Random-effect dose-response model assuming linearity
First we assume a log-linear relation between red meat consumption and diabetes risk using a random-effect model.
```{r results='hold'}
lin <- dosresmeta(formula = logrr ~ dose, type = type, id = id,
	se = se, cases = cases, n = peryears,  data = redmeat)
summary(lin)
```

We found a significant log-linear dose-response association between red meat consumption and diabetes risk (p < 0.0278). There is an evidence of heterogeneity across studies (Q = 75.91, p value = 0.0000).
The change in diabetes risk associated with every 50 grams/day (standard portion) can be obtained with the predict function.
```{r results='hold'}
predict(lin, delta = 50, exp = TRUE)
```
Every 50 grams/day increase in red meat consumption was associated with a significant 12%
(95% CI = 1.2, 24.6) increased risk of diabetes.


## Non-linearity (spline) using random-effect
```{r results='hold'}
library("rms")
knots <- quantile(redmeat$dose, c(.1, .5, .9))
spl <- dosresmeta(formula = logrr ~ rcs(dose, knots), type = type, id = id, se = se, cases = cases, n = peryears, data = redmeat)
summary(spl)
```
## Multivariate Wald test
```{r results='hold'}
library(aod)
wald.test(b = coef(spl), Sigma = vcov(spl), Terms = 1:2)
```
## Tabulate result
```{r results='hold'}
dataTab <- data.frame(dose = seq(0, 200, 50))
predLin <- predict(lin, dataTab, exp = TRUE)
predSpl <- predict(spl, dataTab, exp = TRUE)
round(cbind(lin = predLin, spl = predSpl[2:5]), 2)
```
## figure
```{r results='hold'}
newdata <- data.frame(dose <- seq(0, 200, 1))
xref <- 0
with(predict(spl, newdata, xref, exp = TRUE),{
plot(get("rcs(dose, knots)dose"), pred, type = "l", ylim = c(.8, 1.8),
ylab = "Relative risk", xlab = "Red meat consumption, grams/day",
log = "y", bty = "l", las = 1)
matlines(get("rcs(dose, knots)dose"), cbind(ci.ub, ci.lb),
col = 1, lty = "dashed")
})
points(dose, predict(lin, newdata, xref)$pred, type = "l", lty = 3)
rug(redmeat$dose)
```

Figure 1: Pooled dose-response association between red meat consumption and diabetes risk (solid line). Red meat consumption was modeled with restricted cubic splines in a multivariate random-effects dose-response model. Dash lines represent the 95% confidence intervals for the spline model. The dotted line represents the linear trend. Tick marks below the curve represent the positions of the study-specific relative risks. 
The value of 0 grams/day served as referent. The relative risks are plotted on the log scale.

# Extract Function
```{r results='hold'}
rr_fun <- rcsplineFunction(attr(spl$model[[2]], "parms"), coef(spl))
exp(rr_fun(0))
exp(rr_fun(50))
exp(rr_fun(75))
```

