---
title: "playing with data visualization using ggplot2 in R"
author: "Reynaldi Ikhsan Kosasih"
date: "2023-01-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goals

In this mock-up example, I am going to demonstrate my capability in using R for data visualization

This demonstration used an exam score dataset that consists of 1000 observations.

## load library

These are my 'default' packages when working with r
```{r results='hide'}
library(ggplot2) #create advanced and sophisticating plot
library(tidyverse) #tidying columns, row, etc
library(flextable) #flextable are designed to make tabular reporting easier for R users
library(summarytools) #summary of basic statistics
library(dplyr) #manipulate dataframe
library(readxl) #read excel file
library(writexl) #export to excel

#create a value consists of 25 distinct colors for better plots customization and presentation
c25 <- c(
  "blue1", "#E31A1C", # red
  "green4",
  "#6A3D9A", # purple
  "#FF7F00", # orange
  "black", "gold1",
  "skyblue2", "#FB9A99", # lt pink
  "palegreen2",
  "#CAB2D6", # lt purple
  "#FDBF6F", # lt orange
  "gray70", "khaki2",
  "maroon", "orchid1", "deeppink1", "dodgerblue2", "steelblue4",
  "darkturquoise", "green1", "yellow4", "yellow3",
  "darkorange4", "brown"
)

# load dataset


df <- read_excel("D:/Portfolio/clone/exams score.xlsx", 
     col_types = c("text", "text", "text", 
         "text", "text", "text", "numeric", 
         "numeric", "numeric"))
head(df)
```
## create several new variables as variables of interest

I added several variables, which are overall scores (math * .60 + reading * 0.20 + writing * 0.20) and pass (yes or no). Those who scored more than the minimum passing grade of 75 would pass the exam.
```{r}
df$overall_scores <- (df$math_score * 0.60) + (df$reading_score * 0.20) + (df$writing_score * 0.20)
df$pass <- if_else(df$overall_scores >= 75, "yes", "no", missing = NULL)
```
## check descriptive analysis (frequency, distribution, etc)

```{r}
freq(df$gender)
freq(df$race)
freq(df$parental_education)
freq(df$lunch)
freq(df$test_preparation)
freq(df$pass)

descr(df$math_score)
descr(df$reading_score)
descr(df$writing_score)
descr(df$overall_scores)
```

## using ggplot package to create plots (histogram to observe the distribution)
```{r}
p1 <- ggplot(df, aes(x = overall_scores)) + geom_histogram(binwidth = 3, color="white", fill="lightblue") + geom_vline(aes(xintercept=75), color="maroon", linetype="dashed", size=1) + ggtitle("overall scores of 1000 students") + xlab("overall scores") + ylab("number of students") + scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20)) # adding vertical line of passing grade (75)

p1

# export jpeg file

jpeg(file="histogram overall score.jpeg")
p1
dev.off()
```
## overlaid histogram to compare the overall score between men and women
```{r}
p2 <- ggplot(data=subset(df, !is.na(gender)), aes(x = overall_scores, color = gender)) + geom_histogram(binwidth = 3, fill="white", alpha=0.2, position = "identity") + geom_vline(aes(xintercept=75), color="maroon", linetype="dashed", size=1) + ggtitle("overall scores of 1000 students, men and women") + xlab("overall scores") + ylab("number of students") + scale_y_continuous(limits = c(0, 60), breaks = seq(0, 60, by = 10)) + scale_color_manual(values = c25)

p2
```

## scatterplot for math score, reading score, and writing score
```{r}
# Plot with main and axis titles
library(car)
scatterplot(writing_score ~ reading_score, data = df, grid = TRUE, frame = TRUE, main = "Scatterplot of Reading Score and Writing Score", xlab = "Reading Score", ylab = "Writing Score")

# 3d scatterplot

library(scatterplot3d)

x <- df$math_score
y <- df$reading_score
z <- df$writing_score
grp <- as.factor(df$gender)

scatterplot3d(x, y, z, pch = 16, color = c25[grp],
              grid = TRUE, box = FALSE, xlab = "Math Score", 
              ylab = "Reading Score", zlab = "Writing Score")
legend("right", legend = levels(grp),
      col =  c25[grp], pch = 16)
```
## box plot by race and gender to observe outlier (if any, indicated by red dots)
```{r}
bp <- ggplot(data=subset(df, !is.na(race) & !is.na(gender)), aes(x=race, y=overall_scores, fill=gender)) + geom_boxplot(outlier.colour="red", outlier.shape=16,outlier.size=2, notch=FALSE) + scale_color_manual(values=c25)
bp
```